---
title: "Anomaly detection in Yahoo time series"
output: 
  # html_document
  rmdformats::readthedown:
    highlight: kate
---

# Required libraries

Tidyverse - core tidy verse packges used in this project are:  

* ggplo2 - plotting library, duh
* purr - functional programming toolkit, helps you forget what loops are
* readr - reading rectangular data, .csv in our case
* dplyr - it's like sql, but in r

AnomalyDetection - Twitter's outlier detection package that implements Season Hybrid ESD (Extreme Studentized Deviant).

Forecast - methods and tools for analysing univariate time series.

Tseries - time series analysis and computational finance.

Yardstick - tools for quntifying model performance.

```{r, setup, include=FALSE, echo=FALSE, cache=FALSE}
# Rendering libraries
library(knitr)
# library(rmdformats)

## Global options
options(max.print="75")
opts_chunk$set(cache=TRUE,
               prompt=FALSE,
               # tidy=TRUE,
               comment=NA,
               message=FALSE,
               warning=FALSE)
opts_knit$set(width=75)

# Project libraries
library(tidyverse)
library(AnomalyDetection)
library(forecast)
library(yardstick)
library(tseries)
library(TSstudio)
```

# Loading dataset

Default path is set to `dataset/A1Bechmark`. You can specify whether to convert timestapms to `Date` type.

```{r, cache=TRUE}
load_dataset <- function(convert_timestamp = FALSE,
                         path = "dataset/A1Benchmark") {
  dataset_files <- list.files(path, full.names = TRUE)
  
  # We throw warning if dataset is empty
  if (length(dataset_files) == 0){
    warning(paste("Warning: loaded dataset has 0 records.\n"))
  }
  
  # Parse files
  dataset <- dataset_files %>% 
    map(function(x) {
        read_csv(x, col_type = cols(
          col_double(),
          col_double(),
          col_double()
        ))
      })
  if (convert_timestamp) {
    dataset <- dataset %>%
      map(function(x) {
        time <- as.POSIXlt.Date(x$timestamp)
        df <- data.frame(time, x$value, x$is_anomaly)
        colnames(df) <- c("timestamp", "value", "is_anomaly")
        df
      })
  }
  dataset
}


# Load dataset
dataset <- load_dataset(TRUE)
paste("Yahoo time series dataset contains", length(dataset), "records")
```

# Exploring yahoo dataset

Dataset snippet

```{r}
head(load_dataset()[[30]])
```


Plot time series with highlighted anomalies.

```{r, warning=FALSE}

plot_ts <- function(series) {
  # Create a copy of timse series with normal values set to NA
  outliers_NA <- series
  outliers_NA$value[which(outliers_NA$is_anomaly == 0)] <- NA
  
  # ggplot default color:
  # - bluish green: #00BFC4
  # - vermillion: #F8766D
  ggplot(series, aes(x = timestamp, y = value)) +
    geom_line(size = 0.3) +
    geom_point(outliers_NA, mapping = aes(x = timestamp, y = value, color = "red"),
               size = 3, shape = 1) +
    scale_x_datetime(expand = c(0.01, 0)) +
    scale_y_continuous(expand = c(0.05, 0)) +
    theme_bw() +
    theme(legend.position = "none") +
    labs(x = "Timestamp", y = "Value", color = "Anomaly")
}

# Set working time series
series <- dataset[[30]]
plot_ts(series)
```

Let's add some histograms.

```{r}
plot_hist <- function(series, binsize = 1) {
  ggplot(series, aes(x = value)) +
    geom_histogram(binwidth = binsize, fill = "white", color = "black") +
    theme_bw() +
    labs(x = "Value", y = "Frequency", color = "Anomaly")
}

# Histogram of untrasformed data
plot_hist(series, 0.1)

# Difference the data
diffs <- data.frame(series %>% map(diff))
plot_hist(diffs, 0.1)
```

<!-- Decomposition of the time series. -->

<!-- ```{r} -->
<!-- stl(series$value) -->
<!-- ``` -->

Test if time series is stationary.

```{r, warning=FALSE}
adf.test(series$value)
```

Test presence of autocorrelation.

```{r}
Box.test(series$value)
```

Plotting the autocorrelation function.

```{r}
acf(series$value)
```

Plotting the partial autocorrelation function.

```{r}
pacf(series$value)
```

# Anomaly detection

## Liner regression

### Predicting

`TODO`

### Identifying anomalies

`TODO`

## Seasonal Hybrid ESD model

Find and plot anomalies using Hybrid ESD.

```{r, cache=TRUE, warning=FALSE, message=FALSE}
find_anomalies <- function(series) {
  # Leave only timestamp and value columns
  data = data.frame(series$timestamp, series$value)
  
  # Find and plot anomalies
  res = AnomalyDetectionTs(data,
                           max_anoms = 0.02,
                           direction = "both",
                           plot = TRUE)
  res$plot <- 
    res$plot + 
    geom_line(size = 0.3, color = "black") +
    scale_color_manual(values = c("black", "red")) +
    theme_bw() + 
    theme(legend.position = "none") +
    scale_x_datetime(expand = c(0.01, 0)) +
    scale_y_continuous(expand = c(0.05, 0)) +
    labs(x = "Timestamp", y = "Value", color = "Anomaly")
  res
}

# Plot actual anomalies
plot_ts(series) + ggtitle("Actual anomalies")

# Plot Hybrid ESD anomalies
res = find_anomalies(series)
res$plot + ggtitle("Hybrid ESD anomalies")
```

Plot confusion matrix.

```{r, cache=TRUE}
plot_confusion <- function(series, predicted) {
  real_series <- series$is_anomaly
  
  # Fill prediction series
  predicted_series <- vector("integer", length(real_series))
  predicted_series[match(as.numeric(predicted$anoms$timestamp),
                         as.numeric(series$timestamp))] <- 1
  
  anomalies_df <-
    data.frame(as.factor(real_series), as.factor(predicted_series))
  colnames(anomalies_df) <- c("real", "predicted")
  cm <- conf_mat(anomalies_df, real, predicted)
  autoplot(cm, type = "heatmap")
}

plot_confusion(series, res)
```

## ARIMA model

### Predicting anomalies

Fitting an ARIMA Model.

```{r}
yahoo_m <- auto.arima(series$value, seasonal = FALSE)
yahoo_m
```

Running diagnosis on an ARIMA model.

```{r}
checkresiduals(yahoo_m)
```

Make forecast with ARIMA model.

```{r}
fc_yahoo_m <- forecast(yahoo_m, 20)
fc_yahoo_m
```

Plotting the forecast.

```{r}
ts_length = length(fc_yahoo_m$x)

# Plot ARIMA forecast
autoplot(fc_yahoo_m) +
  theme_bw()

# Zoomed-in version
autoplot(fc_yahoo_m) +
  theme_bw() +
  coord_cartesian(xlim = c(max(ts_length - 100, 0), ts_length))
```

### Identifying anomalies

`TODO`

## Machine learning models

### Random Forests model

`TODO`

### Gradient Boosted Machine model

`TODO`