---
title: "Anomaly detection in Yahoo time series"
output: 
  # html_document
  rmdformats::readthedown:
    highlight: kate
---

# Required libraries

Tidyverse - core tidy verse packges used in this project are:  

* ggplo2 - plotting library, duh
* purr - functional programming toolkit, helps you forget what loops are
* readr - reading rectangular data, .csv in our case
* dplyr - it's like sql, but in r

AnomalyDetection - Twitter's outlier detection package that implements Seasonal Hybrid ESD (Extreme Studentized Deviant).

Forecast - methods and tools for analysing univariate time series.

Tseries - time series analysis

Yardstick - tools for quntifying model performance.

e1071 - Support Vector Machine implementation.

Solitude - isolation forests implementation.

```{r, setup, include=FALSE, echo=FALSE, cache=FALSE}
# Project libraries
# Install, if some of them missing by running: 
#   install.packages('<name of the package>')
library(tidyverse)
library(AnomalyDetection)
library(solitude)
library(forecast)
library(yardstick)
library(tseries)
library(e1071)
```

# Loading dataset

Default path is set to `dataset/`.

```{r, cache=TRUE}
load_dataset <- function(convert_timestamp = FALSE,
                         path = "data") {
  dataset_files <- list.files(path, full.names = TRUE)[1]
  
  # We throw warning if dataset is empty
  if (length(dataset_files) == 0){
    warning(paste("Warning: loaded dataset has 0 records.\n"))
  }
  
  # Parse files
  dataset <- dataset_files %>% 
    map(function(x) {
        read_csv(x, col_type = cols(
          col_double(),
          col_double(),
          col_double()
        ))
      })
  if (convert_timestamp) {
    dataset <- dataset %>%
      map(function(x) {
        time <- as.POSIXlt.Date(x$timestamp)
        df <- data.frame(time, x$value, x$is_anomaly)
        colnames(df) <- c("timestamp", "value", "is_anomaly")
        df
      })
  }
  dataset
}

# Load dataset
dataset <- load_dataset(TRUE)
paste("Yahoo time series dataset contains", length(dataset), "records")
```

# Exploring yahoo dataset

Dataset snippet
```{r}
head(load_dataset()[[1]])
```

Plot time series with highlighted anomalies.

```{r, warning=FALSE}

plot_ts <- function(series, title="") {
  # Create a copy of timse series with normal values set to NA
  series$timestamp <- as.numeric(series$timestamp)
  outliers_NA <- series
  outliers_NA$value[which(outliers_NA$is_anomaly == 0)] <- NA
  
  # ggplot default color:
  # - bluish green: #00BFC4
  # - vermillion: #F8766D
  ggplot(series, aes(x = timestamp, y = value)) +
    ggtitle(title) +
    geom_line(size = 0.3, color = "deepskyblue3") +
    geom_point(outliers_NA, mapping = aes(x = timestamp, y = value, color = "red"),
               size = 3, shape = 1) +
    scale_x_continuous(expand = c(0.01, 0)) +
    scale_x_reverse() +
    scale_y_continuous(expand = c(0.05, 0)) +
    theme_bw() +
    theme(axis.text.x = element_blank()) +
    theme(legend.position = "none") +
    labs(x = "Timestamp", y = "Value", color = "Anomaly")
}

normalize <- function(x) {
  x_std <- (x - min(x)) / max(x) - min(x)
}

# Set working time series
series <- dataset[[1]]
series$value <- normalize(series$value)
plot_ts(series)
```


Let's add some histograms
```{r}
plot_hist <- function(series, binsize = 1) {
  ggplot(series, aes(x = value)) +
    geom_histogram(binwidth = binsize, fill = "white", color = "black") +
    theme_bw() +
    labs(x = "Value", y = "Frequency", color = "Anomaly")
}

# Histogram of untrasformed data
plot_hist(series, 0.02)

# Difference the data
diffs <- data.frame(series %>% map(diff))
plot_hist(diffs, 0.02)
```

<!-- Decomposition of the time series. -->

<!-- ```{r} -->
<!-- stl(series$value) -->
<!-- ``` -->

Test if time series is stationary.

```{r, warning=FALSE}
adf.test(series$value)
```

Test presence of autocorrelation.

```{r}
Box.test(series$value)
```

Plotting the autocorrelation function.

```{r}
acf(series$value)
```

Plotting the partial autocorrelation function.

```{r}
pacf(series$value)
```

# Anomaly detection

## Statistical approach

We can identify anomalies using interquartile distance
```{r, warning=FALSE}
anomalies_STAT <- function(series, threshold = 1.5) {
  df_stat_anom <- series %>%
  select(value) %>%
  mutate(is_low_percentile = case_when(
    value < quantile(value, probs = 0.25) - threshold * IQR(value) |
      value > quantile(value, probs = 0.75) + threshold * IQR(value)
    ~ 1,
    TRUE ~ 0
  ))
  df_stat_anom$is_low_percentile
}

series_stat <- series
series_stat$is_anomaly <- 0
series_stat["is_anomaly"] <- anomalies_STAT(series)
plot_ts(series_stat, "Statistical approach")

```

Plot confusion matrix
```{r}
plot_conf_mat <- function(truth, prediction) {
  labels = data.frame("truth" = as.factor(truth), 
                      "prediction" = as.factor(prediction))
  labels %>% conf_mat(truth, prediction) %>% autoplot(type = "heatmap")
}

plot_conf_mat(series$is_anomaly, anomalies_STAT(series))
```

Calculate precision
```{r}
get_precision <- function(truth, prediction) {
      labels = data.frame("truth" = as.factor(truth), 
                      "prediction" = as.factor(prediction))
    labels %>% precision(truth, prediction)
}

get_precision(series$is_anomaly, series_stat$is_anomaly) %>% print()
```

Calculate recall
```{r}
get_recall <- function(truth, prediction) {
    labels = data.frame("truth" = as.factor(truth), 
                      "prediction" = as.factor(prediction))
    labels %>% recall(truth, prediction)
}

get_recall(series$is_anomaly, anomalies_STAT(series)) %>% print()
```

Calculate F1 score
```{r}
get_f_score <- function(truth, prediction) {
    labels = data.frame("truth" = as.factor(truth), 
                      "prediction" = as.factor(prediction))
    labels %>% f_meas(truth, prediction)
}

get_f_score(series$is_anomaly, anomalies_STAT(series)) %>% print()
```

Calculate precision and recall based on the threashold.
Usually this is done by shifting probability/scores threshold, thus lowering/increasing precision and recall values. In IQR method there is no probability/scores of the value being identified as an anomaly. We still try to find what are precision/recall values based on IQR coefficient threashold. We aren't sure if this method of plotting P-R curve is correct.
```{r}
pr_curve_stat <- function(series, dt = 0.1) {
  
  # Initialize empty precision and recall vector values
  precision_values <- c(1)
  recall_values <- c(0)
  
  # Set cutoff threashold and precision to 0 
  curr_recall <- 0
  curr_threshold <- 0
  
  # Iterate until recall equals 1
  while(curr_recall < 1){
    
    # Update threshold
    curr_threshold <- curr_threshold + dt
    
    # Find anomalies
    predicted_anomalies <- anomalies_STAT(series, curr_threshold)
    
    # Calculate precision and recall
    curr_precision <- get_precision(series$is_anomaly, predicted_anomalies)[[3]]
    curr_recall <- get_recall(series$is_anomaly, predicted_anomalies)[[3]]
    
    # Append current precision and recall values
    precision_values[length(precision_values) + 1] <- curr_precision
    recall_values[length(recall_values) + 1] <- curr_recall
    # print(paste("Precision: ", curr_precision))
    # print(paste("Recall: ", curr_recall))
    # print(" === ")
  }
  return (data.frame(precision = precision_values,
                   recall = recall_values))
}

pr_values <- pr_curve_stat(series, 0.01)
```

Plot precision-recall curve
```{r}
ggplot(pr_values, aes(x = recall, y = precision)) + 
  geom_line()
```


## One Class SVM

Grid search
```{r}
# Grid search parameters
gs <- list(window = c(5, 7, 10),
           nu = c(0.05, 0.01, 0.1, 0.2),
           kernel = c("radial", "sigmoid", "polynomial", "linear")) %>% cross_df

gs <- data.frame(gs)

# Intialize and train model based on provided hyperparameters
# Return corresponding outliers
ocsmv_outliers <- function(series, window, nu, kernel){
  
  # Embedding dimensions of the time series
  series_embeddings <- embed(series$value, window)
  
  # Create model based on the paramters
  model_ocsmv <- svm(series_embeddings, 
                   type = 'one-classification', 
                   kernel = kernel,
                   gamma = 0.05,
                   nu = nu)
  
  # Train model
  pred_ocsvm <- predict(model_ocsmv, series_embeddings)
  
  # Store detected outliers in new time series
  series_ocsvm <- series
  series_ocsvm$is_anomaly <- 0
  for (idx in which(pred_ocsvm == FALSE)){
    series_ocsvm$is_anomaly[idx:idx+4] = 1
  }
  series_ocsvm
}

gs$scores <- 0
for (idx in 1:nrow(gs)){
  outliers <- ocsmv_outliers(series, gs[idx, ]$window, gs[idx, ]$nu, gs[idx, ]$kernel)
  gs$scores[idx] <- get_f_score(series$is_anomaly, outliers$is_anomaly)[[3]]
}

print(gs[order(gs$scores, decreasing = TRUE), ])
```


Embeding time series values
```{r}
series_embeddings <- embed(series$value, 5)
data.frame(t(series_embeddings[1, ]))
```

```{r}
model_ocsmv <- svm(series_embeddings, 
                   type = 'one-classification', 
                   kernel = 'radial',
                   gamma = 0.05,
                   nu = 0.01)
model_ocsmv
```

Identifying outliers using above model
```{r}
pred_ocsvm <- predict(model_ocsmv, series_embeddings)

series_ocsvm <- series
series_ocsvm$is_anomaly <- 0
for (idx in which(pred_ocsvm == FALSE)){
  series_ocsvm$is_anomaly[idx:idx+4] = 1
}
plot_ts(series_ocsvm, "One-class SVM anomalies")
```

Confusion matrix
```{r}
plot_conf_mat(series$is_anomaly, series_ocsvm$is_anomaly)
```

Precisions
```{r}
get_precision(series$is_anomaly, series_ocsvm$is_anomaly) %>% print()
```

Recall
```{r}
get_recall(series$is_anomaly, series_ocsvm$is_anomaly) %>% print()
```

F1 score
```{r}
get_f_score(series$is_anomaly, series_ocsvm$is_anomaly) %>% print()
```

## Seasonal Hybrid ESD model

Find and plot anomalies using Hybrid ESD.

```{r, cache=TRUE, warning=FALSE, message=FALSE}
anomalies_ESD <- function(series) {
  # Leave only timestamp and value columns
  data = data.frame(series$timestamp, series$value)

  # Find and plot anomalies
  res = AnomalyDetectionTs(data,
                           max_anoms = 0.02,
                           direction = "both",
                           plot = TRUE)
  series_esd <- series
  series_esd$is_anomaly <- 0
  series_esd$is_anomaly[as.numeric(series$timestamp) %in% as.numeric(res$anoms$timestamp)] <- 1
  return(series_esd)
}

# Plot Hybrid ESD anomalies
anoms_esd = anomalies_ESD(series)
plot_ts(anoms_esd, "Seasonal Hybrid ESD anomalies")
```

Confusion matrix
```{r}
plot_conf_mat(series$is_anomaly, anoms_esd$is_anomaly)
```

Precisions
```{r}
get_precision(series$is_anomaly, anoms_esd$is_anomaly) %>% print()
```

Recall
```{r}
get_recall(series$is_anomaly, anoms_esd$is_anomaly) %>% print()
```

F1 score
```{r}
get_f_score(series$is_anomaly, anoms_esd$is_anomaly) %>% print()
```


## ARIMA model

### Predicting anomalies

Fitting an ARIMA Model.

```{r}
yahoo_m <- auto.arima(series$value, seasonal = FALSE)
yahoo_m
```

Running diagnosis on an ARIMA model.

```{r}
checkresiduals(yahoo_m)
```

Make forecast with ARIMA model.

```{r}
fc_yahoo_m <- forecast(yahoo_m, 20)
fc_yahoo_m
```

Plotting the forecast.

```{r}
ts_length = length(fc_yahoo_m$x)

# Plot ARIMA forecast
autoplot(fc_yahoo_m) +
  theme_bw()
```

### Identifying anomalies

For some reason ARIMA model can't forecast time series value beyound 10 steps. As seen on forecasting plot the forecast in smoothed out.

## Machine learning models

### Isolation forests

Detecting anomalies with isoaltion forests using `solitude` library
```{r, warning=FALSE, message=FALSE}

anomalies_ISO <- function(series, threashold=0.5) {
  # Isolation forests model instance
  iso <- isolationForest$new()
  
  # Fit for time series values data
  iso$fit(data.frame(series$value))
  
  # Identify anomalies score above the threashold
  iso_anomalies <- if_else(iso$scores$anomaly_score > threashold, 1, 0)
}

anomalies_ISO_scores <- function(series){
    # Isolation forests model instance
  iso <- isolationForest$new()
  
  # Fit for time series values data
  iso$fit(data.frame(series$value))
  data.frame(truth = as.factor(series$is_anomaly),
             estimate = 1 - iso$scores$anomaly_score)
}

```

```{r}
scores_iso <- anomalies_ISO_scores(series)
pr_curve(scores_iso, truth, estimate) %>% autoplot()
```


```{r}
threashold = 0.6
series_iso <- series
series_iso["is_anomaly"] <- anomalies_ISO(series, threashold)
plot_ts(series_iso, paste("Isolation forest anomalies. Threashold:", threashold))
```

Confusion matrix
```{r}
plot_conf_mat(series$is_anomaly, series_iso$is_anomaly)
```

Precisions
```{r}
get_precision(series$is_anomaly, series_iso$is_anomaly) %>% print()
```

Recall
```{r}
get_recall(series$is_anomaly, series_iso$is_anomaly) %>% print()
```

F1 score
```{r}
get_f_score(series$is_anomaly, series_iso$is_anomaly) %>% print()
```

